Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
edits to 2.1

We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
The recognition model exploits the structure of the DSL: %$\mathcal{D}$:
rather than directly predicting a distribution over $p$ conditioned on $x$,
it predicts a distribution over components of the DSL.
This taps into the intuition that programming is primarily a top-down activity:
as human programmers, we often first decide what kind of
programming constructs we might need to use,
and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

>>>>

We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
  \nonumber
    J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
  \nonumber
    \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most likely increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
Rather than directly predicting a distribution over $p$ conditioned on $x$,
the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
This taps into the intuition that programming is primarily a top-down activity:
as human programmers, we often first decide what kind of
programming constructs we might need to use,
and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.
